\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-10-14 papers/methods/paper.tex
emulateapj

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{mnras}

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[\name]{\name: A hierarchical Bayesian model for Supernova Cosmology}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[S. R. Hinton et al.]{
	Samuel R. Hinton,$^{1,2}$\thanks{E-mail: samuelreay@gmail.com}
	Alex G. Kim,$^{3}$
	Tamara M. Davis$^{1,2}$
	\\
	% List of institutions
	$^{1}$School of Mathematics and Physics, The University of Queensland, Brisbane, QLD 4072, Australia\\
	$^{2}$ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)\\
	$^{3}$Physics Division, Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2017}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
We present a new Bayesian hierarchical model named {\steve} for performing type Ia supernova (SN~Ia) cosmology fits. 
%
This advances previous works with Bayesian hierarchical models by including 
an improved treatment of Malmquist bias, 
accounting for additional sources of systematic uncertainty, 
and increasing numerical efficiency. 
%
Given light curve fit parameters, redshifts, and host-galaxy masses, {\green we fit {\steve}} simultaneously for parameters describing cosmology, SN~Ia populations, and systematic uncertainties. Selection effects are characterised using Monte-Carlo simulations.
%
We demonstrate its implementation by fitting realisations of SN~Ia datasets where the SN~Ia model closely follows that used in {\steve}.
%
Next, we validate on more realistic SNANA simulations of SN~Ia samples from the Dark Energy Survey and low-redshift surveys \citep{DESKEY}.
%
These simulated datasets contain more than $60\,000$ SNe~Ia, which we use to evaluate biases in the recovery of cosmological parameters, specifically the equation-of-state of dark energy, $w$. 
%
This is the most rigorous test of a BHM model, and reveals small $w$-biases that depend on the simulated SN~Ia properties, in particular the intrinsic SN~Ia scatter model. This $w$-bias is less than $0.03$ on average, less than half the statistical uncertainty on $w$.
%
These simulation test results are a concern {\green for BHM applications on large upcoming surveys}, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
\end{abstract}
% Bias in SNANA tests across intrinsic dispersion models is found to range from $-0.15\sigma$ to $+0.27\sigma$ in $w$ for datasets of comparable strength to the Dark Energy Survey three-year analysis.

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

@@

Two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}. Since that time, the number of observed type Ia supernovae (SN~Ia) has increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, chambers2016panstarrs, sako2018sdss}. Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complementary probes of large scale structure and the CMB. For a recent review, see \citet{Huterer2018}. {\green While these efforts have reduced the uncertainty on the equation-of-state of dark energy ($w$) by more than a factor of two, it is still consistent with a cosmological constant and the nature of dark energy remains an unsolved mystery.}

In attempts to tease out the nature of dark energy, active and planned surveys are {\green continually growing in size and scale}. The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has {\green discovered} thousands of type Ia supernovae, attaining both spectroscopically and photometrically identified samples. The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will {\green discover tens} of thousands of photometrically classified supernovae. {\green Such increased statistical power demands greater fidelity and flexibility in modelling supernovae for cosmological purposes, as we will require reduced systematic uncertainties to fully utilise these increased statistics} \citep{Betoule2014, Scolnic2017}.

As such, {\green considerable resources are aimed at} developing more sophisticated supernova cosmology analyses. The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific supernova models. First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved for {\green SDSS \citep{Kessler2009},} simulations are a fundamental component of modern supernova cosmology.  \citet{Betoule2014} quantise and correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN~Ia distances as a function of {\green multiple factors to improve bias correction}. Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}. Bayesian Hierarchical models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, {\green and either use simulation-determined distance-corrections to correct for biases, or attempt to find analytic approximations for effects such as Malmquist bias to model the biases inside the BHM itself}.

{\green In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015}. We include additional sources of systematic uncertainty, including an analytic formulation of selection efficiency which incorporates parametric uncertainty. We also implement a different model of intrinsic dispersion to both incorporate redshift-dependent scatter and to increase numerical efficiency, allowing our model to perform rapid fits to supernovae datasets.}

Section \ref{sec:review} is dedicated to a quick review of {\green supernova cosmology analysis methods}, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods. In Section \ref{sec:method} we outline our methodology. Model verification on simulated datasets is given in Section \ref{sec:verification}, along with details on potential areas of improvement. We summarise our methodology in Section \ref{sec:conclusion}.

@@

Whilst supernova observations take the form of photometric time-series brightness measurements in many bands {\green and a redshift measurement of the supernova (or its assumed host)}, most analyses do not work from these measurements directly. Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernova light curves for cosmology \citep{Guy2010, Betoule2014}. The resulting output from the model is, for each supernova, an amplitude $x_0$ (which can be converted into apparent magnitude, $m_B = -2.5\log(x_0)$), a stretch term $x_1$, and color term $c$, along with a covariance matrix describing the uncertainty on these summary statistics ($\mathcal{C_\eta}$). As all supernovae are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.

This {\green ensemble of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$} represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population. Accurately characterising this underlying population, its evolution over redshift, and effects from {\green host-galaxy} environment, is one of the challenges of supernova cosmology. Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ {\green(absolute magnitude of the supernova, traditionally in the Bessell $B$ band)}, $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population from absolute magnitude space into the observed population in apparent magnitude space. Specifically, for any given supernova our map between {\green absolute magnitude and apparent magnitude is given by the distance modulus:}

\begin{align}
\mu_{\rm obs} = m_B + \alpha x_1 - \beta c - M_B + \text{other corrections}, \label{eq:standard}
\end{align}

where $M_B$ is the mean absolute magnitude for all SN~Ia given $x_1=c=0$,  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the color correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter. {\green The distance modulus $\mu_{\rm obs}$ is a product of our observations, however a distance modulus $\mu_C$ can be precisely calculated given cosmological parameters and a redshift.} The `$\text{other corrections}$' term often includes corrections for host galaxy environment, as this has statistically significant correlations with supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}. In traditional $\chi^2$ analyses, Malmquist bias corrections are also included, which can take the form of a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Kessler2017, Scolnic2017}.

@@

Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from equation \ref{eq:standard}, and the cosmologically predicted distance modulus. {\green The $\chi^2$ function minimised is}

@@

where $C^{-1}_{\mu}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Brout18SYS} for a review of these uncertainties for the DES supernova analysis). The predicted $\mu_C$ is defined as
{\green 
\begin{align}
\mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
E(z) &= \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
\end{align}
}
where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$, and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.

The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with several limitations. Firstly, formulating a $\chi^2$ likelihood results in a loss of model flexibility by assuming Gaussian uncertainty. Secondly, the method of creating a covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty. For example, the underlying supernova color population's mean and skewness are highly correlated, however this correlation is ignored when determining population uncertainty using numerical derivatives of population permutations. {\green Whilst correlations can be incorporated into a covariance matrix, it requires human awareness of the correlations and thus methods which can automatically capture correlated uncertainties are preferable.} Thirdly, the computational efficiency is dependent on both creating the global covariance matrix, and then inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae. A recent solution to this computational cost problem is to bin the supernovae {\green in redshift bins}, which produces a matrix of manageable size {\green and can allow fitting without matrix inversion on every step}. Whilst binning data results in some loss of information, recent works tested against simulations show that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.

Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting the determined $\mu_{\rm obs}$ from the data, or equivalently, adding a distance bias to the $\mu_C$ prediction. Specifically, Malmquist bias is the result of losing the fainter tail of the supernova population at high redshift. An example of Malmquist bias is illustrated in Figure \ref{fig:malmquist}, which simulates supernovae according to equation \eqref{eq:standard}. Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then {\green subtracted from} the observational data. When using traditional fitting methods such as that found in \citet{Betoule2014}, these effects are not built into the likelihood and instead are formed by correcting data. This means that the bias uncertainty is not captured fully in the $\chi^2$ distribution, and subtle correlations between cosmological or population parameters and the bias correction is lost. {\green Recent developments such as the BBC method \citep{Kessler2017} incorporate corrections dependent on $\alpha$ and $\beta$, improving their capture of uncertainty on bias corrections in the $\chi^2$ likelihood.}

@@

\subsection{Hierarchical Bayesian Models}

Sitting between the traditional models simplicity and the complexity of forward modelling lies Bayesian hierarchical models (BHM). Hierarchical models utilise multiple layers of connected parameters, with the layers linked via well defined and {\green physically} motivated conditional probabilities. For example, an observation of a parameter from a population will be conditioned on the true value of the parameter, which itself will be conditioned on the population distribution of that parameter. We can thus incorporate different population distributions, and parameter inter-dependence which cannot be found in traditional analyses where uncertainty must be encapsulated in a covariance matrix.

With the introduction of multiple layers in our model, we can add more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood. \citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light-curve fitting. \citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but do not incorporate bias correction. \citet{March2014} and \citet{Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS-like simulations. The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset \citep[using redshift dependent bias corrections from][]{Betoule2014}, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$. \citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}. Notably, these methods rely on data that is bias corrected or {\green the methods }ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}. The assumption made by the UNITY analysis is that the bias {\green in real data} is {\green perfectly described} by an analytic function. {\green They validate their } model to be free of significant biases using fits to {\green thirty realisations of supernova datasets that are constructed to mimic the UNITY framework}. The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, most lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}. For the latter case, by inferring biases using Bayesian models, sophisticated corrections can be calculated and then applied to more traditional $\chi^2$ models.

Whilst there are a large number of hierarchical models available, none of them have undergone high-statistic tests using realistic simulations to verify each models' respective bias. {\green Additionally, testing has generally been performed on supernovae simulations with either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology. However, quantifying the biases on $w$CDM cosmology simulations with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own, and focus shifts from determination of $\Omega_m$ to both $\Omega_m$ and $w$.}

The flexibility afforded by hierarchical models allows for investigations into different treatments of underlying supernova magnitude, color and stretch populations, host-galaxy corrections, and redshift evolution, each of which will be discussed further in the outline of our model below. Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models. {\green We reduce our dependence on an assumed scatter model in simulations by not utilising bias-corrections in an effort} to provide a valuable cross-check on analysis methodologies which utilise scatter-model-dependant bias corrections.

@@

{\green The diverse approaches and implementations applied to supernova cosmology are a response to the significant challenges and complications faced by analyses.} In this Section, we outline several of the most prominent challenges.

{\green Forefront among these challenges is our ignorance of the underlying type Ia intrinsic dispersion. Ideally, analysis of the underlying dispersion would make use of an ensemble of time-series spectroscopy to characterise the diversity of type Ia supernovae. However this data is difficult to obtain, and recent efforts to quantify the dispersion draw inference from photometric measurements. The underlying dispersion model is not a solved problem, and we therefore test against }two dispersion models in this work. The {\green first is based on the} \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation. The second, {\green based on the} {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation. {\green These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.} As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$, and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C_\eta}$. These two scatter models are currently {\green assumed} to span the possible range of scatter in the underlying supernova population. We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{emulateapj}

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.

\shorttitle{\name}
% Don't change these lines
\begin{document}
\title{\name: A hierarchical Bayesian model for Supernova Cosmology}

 The list of authors, and the short list which is used in the headers.
 % If you need two or more lines of authors, add an extra line using \newauthor
 \author{
 	S.~R.~Hinton\altaffilmark{1,2},\thanks{E-mail: samuelreay@gmail.com}
 	T.~M.~Davis\altaffilmark{1,2},
 	A.~G.~Kim\altaffilmark{3},
 	A.~M\"oller\altaffilmark{2,4},
 	M.~Sako\altaffilmark{5},
 	M.~Smith\altaffilmark{6},
 	\\}
 \affil{$^{1}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
 \affil{$^{2}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
 \affil{$^{3}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
 \affil{$^{4}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
 \affil{$^{5}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
 \affil{$^{6}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}

% Abstract of the paper
\begin{abstract}
We present a new Bayesian hierarchical model named {\steve} for performing type Ia supernova (SN~Ia) cosmology fits. 
%
This advances previous works with Bayesian hierarchical models by including 
an improved treatment of Malmquist bias, 
accounting for additional sources of systematic uncertainty, 
and increasing numerical efficiency. 
%
Given light curve fit parameters, redshifts, and host-galaxy masses, we fit {\steve} simultaneously for parameters describing cosmology, SN~Ia populations, and systematic uncertainties. Selection effects are characterised using Monte-Carlo simulations.
%
We demonstrate its implementation by fitting realisations of SN~Ia datasets where the SN~Ia model closely follows that used in {\steve}.
%
Next, we validate on more realistic SNANA simulations of SN~Ia samples from the Dark Energy Survey and low-redshift surveys \citep{DESKEY}.
%
These simulated datasets contain more than $60\,000$ SNe~Ia, which we use to evaluate biases in the recovery of cosmological parameters, specifically the equation-of-state of dark energy, $w$. 
%
This is the most rigorous test of a BHM model, and reveals small $w$-biases that depend on the simulated SN~Ia properties, in particular the intrinsic SN~Ia scatter model. This $w$-bias is less than $0.03$ on average, less than half the statistical uncertainty on $w$.
%
These simulation test results are a concern for BHM applications on large upcoming surveys, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
\end{abstract}

@@

Two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}. Since that time, the number of observed type Ia supernovae (SN~Ia) has increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, chambers2016panstarrs, sako2018sdss}. Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complementary probes of large scale structure and the CMB. For a recent review, see \citet{Huterer2018}. While these efforts have reduced the uncertainty on the equation-of-state of dark energy ($w$) by more than a factor of two, it is still consistent with a cosmological constant and the nature of dark energy remains an unsolved mystery.

In attempts to tease out the nature of dark energy, active and planned surveys are continually growing in size and scale. The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has discovered thousands of type Ia supernovae, attaining both spectroscopically and photometrically identified samples. The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will discover tens of thousands of photometrically classified supernovae. Such increased statistical power demands greater fidelity and flexibility in modelling supernovae for cosmological purposes, as we will require reduced systematic uncertainties to fully utilise these increased statistics \citep{Betoule2014, Scolnic2017}.

As such, considerable resources are aimed at developing more sophisticated supernova cosmology analyses. The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific supernova models. First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved for SDSS \citep{Kessler2009}, simulations are a fundamental component of modern supernova cosmology.  \citet{Betoule2014} quantise and correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN~Ia distances as a function of multiple factors to improve bias correction. Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}. Bayesian Hierarchical models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, and either use simulation-determined distance-corrections to correct for biases, or attempt to find analytic approximations for effects such as Malmquist bias to model the biases inside the BHM itself.

In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015}. We include additional sources of systematic uncertainty, including an analytic formulation of selection efficiency which incorporates parametric uncertainty. We also implement a different model of intrinsic dispersion to both incorporate redshift-dependent scatter and to increase numerical efficiency, allowing our model to perform rapid fits to supernovae datasets.

Section \ref{sec:review} is dedicated to a quick review of supernova cosmology analysis methods, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods. In Section \ref{sec:method} we outline our methodology. Model verification on simulated datasets is given in Section \ref{sec:verification}, along with details on potential areas of improvement. We summarise our methodology in Section \ref{sec:conclusion}.

@@

Whilst supernova observations take the form of photometric time-series brightness measurements in many bands and a redshift measurement of the supernova (or its assumed host), most analyses do not work from these measurements directly. Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernova light curves for cosmology \citep{Guy2010, Betoule2014}. The resulting output from the model is, for each supernova, an amplitude $x_0$ (which can be converted into apparent magnitude, $m_B = -2.5\log(x_0)$), a stretch term $x_1$, and color term $c$, along with a covariance matrix describing the uncertainty on these summary statistics ($\mathcal{C_\eta}$). As all supernovae are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.

This ensemble of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population. Accurately characterising this underlying population, its evolution over redshift, and effects from host-galaxy environment, is one of the challenges of supernova cosmology. Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude of the supernova, traditionally in the Bessell $B$ band), $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population from absolute magnitude space into the observed population in apparent magnitude space. Specifically, for any given supernova our map between absolute magnitude and apparent magnitude is given by the distance modulus:

\begin{align}
\mu_{\rm obs} = m_B + \alpha x_1 - \beta c - M_B + \Delta M \cdot m + \text{other corrections}, \label{eq:standard}
\end{align}

where $M_B$ is the mean absolute magnitude for all SN~Ia given $x_1=c=0$,  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the color correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter. $\Delta M \cdot m$ refers to the host-galaxy mass correlation discussed in Section \ref{sec:hostgal}. The distance modulus $\mu_{\rm obs}$ is a product of our observations, however a distance modulus $\mu_C$ can be precisely calculated given cosmological parameters and a redshift. The `$\text{other corrections}$' term often includes bias corrections for traditional $\chi^2$ analyses. Bias corrections can take multiple forms, such as a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Kessler2017, Scolnic2017}.

@@

Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from equation \ref{eq:standard}, and the cosmologically predicted distance modulus. The $\chi^2$ function minimised is

@@

where $C^{-1}_{\mu}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Brout18SYS} for a review of these uncertainties for the DES supernova analysis). The predicted $\mu_C$ is defined as

\begin{align}
\mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
E(z) &= \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
\end{align}

where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$, and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.

The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with several limitations. Firstly, formulating a $\chi^2$ likelihood results in a loss of model flexibility by assuming Gaussian uncertainty. Secondly, the method of creating a covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty. For example, the underlying supernova color population's mean and skewness are highly correlated, however this correlation is ignored when determining population uncertainty using numerical derivatives of population permutations. Whilst correlations can be incorporated into a covariance matrix, it requires human awareness of the correlations and thus methods which can automatically capture correlated uncertainties are preferable. Thirdly, the computational efficiency is dependent on both creating the global covariance matrix, and then inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae. A recent solution to this computational cost problem is to bin the supernovae in redshift bins, which produces a matrix of manageable size and can allow fitting without matrix inversion on every step. Whilst binning data results in some loss of information, recent works tested against simulations show that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.

Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting the determined $\mu_{\rm obs}$ from the data, or equivalently, adding a distance bias to the $\mu_C$ prediction. Specifically, Malmquist bias is the result of losing the fainter tail of the supernova population at high redshift. An example of Malmquist bias is illustrated in Figure \ref{fig:malmquist}, which simulates supernovae according to equation \eqref{eq:standard}. Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then subtracted from the observational data. When using traditional fitting methods such as that found in \citet{Betoule2014}, these effects are not built into the likelihood and instead are formed by correcting data. This means that the bias uncertainty is not captured fully in the $\chi^2$ distribution, and subtle correlations between cosmological or population parameters and the bias correction is lost. Recent developments such as the BBC method \citep{Kessler2017} incorporate corrections dependent on $\alpha$ and $\beta$, improving their capture of uncertainty on bias corrections in the $\chi^2$ likelihood.

@@

\subsection{Bayesian Hierarchical Models}

Sitting between the traditional models simplicity and the complexity of forward modelling lies Bayesian hierarchical models (BHM). Hierarchical models utilise multiple layers of connected parameters, with the layers linked via well defined and physically motivated conditional probabilities. For example, an observation of a parameter from a population will be conditioned on the true value of the parameter, which itself will be conditioned on the population distribution of that parameter. We can thus incorporate different population distributions, and parameter inter-dependence which cannot be found in traditional analyses where uncertainty must be encapsulated in a covariance matrix. Unlike ABC methods, which can model arbitrary probability distributions, BHM methods are generally constrained to representing probabilities using analytic forms.

With the introduction of multiple layers in our model, we can add more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood. \citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light-curve fitting. \citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but do not incorporate bias correction. \citet{March2014} and \citet{Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS-like simulations. The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset \citep[using redshift dependent bias corrections from][]{Betoule2014}, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$. \citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}. Notably, these methods rely on data that is bias corrected or the methods ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}. The assumption made by the UNITY analysis is that the bias in real data is perfectly described by an analytic function. They validate their  model to be free of significant biases using fits to thirty realisations of supernova datasets that are constructed to mimic the UNITY framework. The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, most lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}. For the latter case, by inferring biases using Bayesian models, sophisticated corrections can be calculated and then applied to more traditional $\chi^2$ models.

Whilst there are a large number of hierarchical models available, none of them have undergone high-statistic tests using realistic simulations to verify each models' respective bias. Additionally, testing has generally been performed on supernovae simulations with either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology. However, quantifying the biases on $w$CDM cosmology simulations with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own, and focus shifts from determination of $\Omega_m$ to both $\Omega_m$ and $w$.

The flexibility afforded by hierarchical models allows for investigations into different treatments of underlying supernova magnitude, color and stretch populations, host-galaxy corrections, and redshift evolution, each of which will be discussed further in the outline of our model below. Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models. We reduce our dependence on an assumed scatter model in simulations by not utilising bias-corrections in an effort to provide a valuable cross-check on analysis methodologies which utilise scatter-model-dependant bias corrections.

@@

 The diverse approaches and implementations applied to supernova cosmology are a response to the significant challenges and complications faced by analyses. In this Section, we outline several of the most prominent challenges.

 Forefront among these challenges is our ignorance of the underlying type Ia intrinsic dispersion. Ideally, analysis of the underlying dispersion would make use of an ensemble of time-series spectroscopy to characterise the diversity of type Ia supernovae. However this data is difficult to obtain, and recent efforts to quantify the dispersion draw inference from photometric measurements. The underlying dispersion model is not a solved problem, and we therefore test against two dispersion models in this work. The first is based on the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, the second is sourced from \citet[][hereafter denoted {\celeven}]{Chotard2011}. As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$, and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C_\eta}$. These two scatter models are currently assumed to span the possible range of scatter in the underlying supernova population. We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.
